Multi-EWAS enrichment example

### 2024-01-16 figs/multi_ewas_example.Rmd
fix color pal
incidence

myfiles <- list.files("hillary",pattern="prevalent_full",full.names=TRUE)
l1 <- lapply(myfiles, read.csv)
l1 <- lapply(l1,function(x) { x[,c("CpG","Beta"),drop=FALSE] } )
names(l1) <- gsub("_gs.csv","",gsub("hillary/prevalent_full_","",myfiles))
str(l1)

@@

@@

tic()
m <- mitch_import(x=l1,DEtype="prescored",geneTable=gt,geneIDcol="CpG")
toc() #117 sec
head(m)

@@

## Mitch calc

Multi-comparison.

```{r,mitch1}

tic()
mres <- mitch_calc(x=m,genesets=gs_symbols,minsetsize=5,cores=8, priority="effect")
toc() #21.3s

mtable <- mres$enrichment_result
head(mtable,20)

#numsig manova
nrow(subset(mtable,p.adjustMANOVA<0.05))
sig <- subset(mtable,p.adjustMANOVA<0.05)

@@

```{r,mitch2,fig.height=5,fig.width=5}

@@

r2 <- lapply(r1,function(r) {
  ups <- head(subset(r,p.adjustANOVA<0.05 & s.dist > 0.3)$set,5)
  dns <- head(subset(r,p.adjustANOVA<0.05 & s.dist < -0.3)$set,5)
  list("ups"=ups,"dns"=dns)
} )

@@

```{r,mitchheat,fig.height=9,fig.width=9}

top <- mtable[which(mtable$set %in% gsets),]

top <- top[,c(1,4:17)]

rownames(top) <- top$set
top$set=NULL
head(top,2)

rownames(top) <- gsub("REACTOME_","",rownames(top))
rownames(top) <- gsub("_"," ",rownames(top))

colnames(top) <- gsub("^s.","",colnames(top))

colnames(top) <- gsub("_"," ",colnames(top))

#colfunc <- colorRampPalette(c("#B80F0A", "white", "#1134A6"))
colfunc <- colorRampPalette(c("red", "white", "blue"))

@@

>>>>

myfiles1 <- list.files("hillary",pattern="prevalent_full",full.names=TRUE)
l1 <- lapply(myfiles1, read.csv)
l1 <- lapply(l1,function(x) { x[,c("CpG","Beta"),drop=FALSE] } )
names(l1) <- gsub("_gs.csv","",gsub("hillary/prevalent_full_","",myfiles))
str(l1)

myfiles2 <- list.files("hillary",pattern="incident_full",full.names=TRUE)
l2 <- lapply(myfiles2, read.csv)
l2 <- lapply(l2,function(x) { x[,c("CpG","Beta"),drop=FALSE] } )
names(l2) <- gsub("_gs.csv","",gsub("hillary/incident_full_","",myfiles2))
str(l2)

@@

l1: prevalent

l2: incident

@@

tic()
m1 <- mitch_import(x=l1,DEtype="prescored",geneTable=gt,geneIDcol="CpG")
toc() #117 sec
head(m1)

tic()
m2 <- mitch_import(x=l2,DEtype="prescored",geneTable=gt,geneIDcol="CpG")
toc() #158 sec
head(m2)

@@

## Mitch calc prevalence

Multi-comparison for prevalence.

```{r,mitch1prev}

tic()
mres1 <- mitch_calc(x=m1,genesets=gs_symbols,minsetsize=5,cores=8, priority="effect")
toc() #21.3s

mtable1 <- mres1$enrichment_result
head(mtable1,20)

#numsig manova
nrow(subset(mtable1,p.adjustMANOVA<0.05))
sig <- subset(mtable1,p.adjustMANOVA<0.05)

@@

```{r,mitch2prev,fig.height=5,fig.width=5}

@@

r2 <- lapply(r2,function(r) {
  ups <- head(subset(r,p.adjustANOVA<0.05 & s.dist > 0.3)$set,5)
  dns <- head(subset(r,p.adjustANOVA<0.05 & s.dist < -0.3)$set,5)
  list("ups"=ups,"dns"=dns)
} )

@@

```{r,mitchheatprev,fig.height=9,fig.width=9}

top <- mtable1[which(mtable1$set %in% gsets),]
top <- top[,c(1,4:17)]
rownames(top) <- top$set
top$set=NULL
head(top,2)

rownames(top) <- gsub("REACTOME_","",rownames(top))
rownames(top) <- gsub("_"," ",rownames(top))
colnames(top) <- gsub("^s.","",colnames(top))
colnames(top) <- gsub("_"," ",colnames(top))
colfunc <- colorRampPalette(c("blue", "white", "red"))

@@

## Mitch calc incidence

Multi-comparison for incidence.

```{r,mitch1incidence}

tic()
mres2 <- mitch_calc(x=m2,genesets=gs_symbols,minsetsize=5,cores=8, priority="effect")
toc() #29.5s

mtable2 <- mres2$enrichment_result
head(mtable2,20)

#numsig manova
nrow(subset(mtable2,p.adjustMANOVA<0.05))
sig <- subset(mtable2,p.adjustMANOVA<0.05)

```{r,mitch2incidence,fig.height=5,fig.width=5}

r1 <- lapply(l2,function(m) {
  m2 <- mitch_import(x=m,DEtype="prescored",geneTable=gt,geneIDcol="CpG")
  mres <- mitch_calc(x=m2,genesets=gs_symbols,minsetsize=5,cores=8, priority="effect")
  mtable <- mres$enrichment_result
})

rownames(nums) <- gsub("ibd","IBD",rownames(nums))
rownames(nums) <- gsub("_"," ",rownames(nums))

pdf("fig8c.pdf",height=6,width=4)
par(mar = c(5.1, 10.1, 4.1, 2.1))
barplot(t(nums),beside=TRUE,horiz=TRUE,las=1,
  legend.text = c("down","up"),xlab="no. pathways")
abline(v=seq(0,200,50),lty=2,lwd=0.5,col="gray")
dev.off()

r2 <- lapply(r1,function(r) {
  ups <- head(subset(r,p.adjustANOVA<0.05 & s.dist > 0.3)$set,3)
  dns <- head(subset(r,p.adjustANOVA<0.05 & s.dist < -0.3)$set,3)
  list("ups"=ups,"dns"=dns)
} )

```{r,mitchheatincidence,fig.height=9,fig.width=9}

top <- mtable2[which(mtable2$set %in% gsets),]
top <- top[,c(1,4:17)]
rownames(top) <- top$set
top$set=NULL
head(top,2)

rownames(top) <- gsub("REACTOME_","",rownames(top))
rownames(top) <- gsub("_"," ",rownames(top))
colnames(top) <- gsub("^s.","",colnames(top))
colnames(top) <- gsub("_"," ",colnames(top))
colnames(top) <- gsub("ibd","IBD",colnames(top))

colfunc <- colorRampPalette(c("blue", "white", "red"))

heatmap.2(as.matrix(top),scale="none",trace="none",margins=c(6,22),
  col=colfunc(25),cexRow=0.6,cexCol=0.7)

pdf("fig8d.pdf",height=9,width=7)
heatmap.2(as.matrix(top),scale="none",trace="none",margins=c(6,20),
  col=colfunc(25),cexRow=0.5,cexCol=0.8)
dev.off()
